---
title: "Companion to Assessing shop completeness in OpenStreetMap for two federalstates in Germany - Query OSM Retail stores in Baden-Württemberg and Saxony"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
```{r setup, include=FALSE}
# packages with need to be installed (once) with install.packages("httr") etc.
require(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(readtext)
library(rapiclient)
require(RCurl)
require(sf)
require(geojsonsf)
#require(geojsonR)
require(tmap)
require(tmaptools)
library(rgdal)
require(geojsonio)
require(geojson)
#require(osmdata)
require(rnaturalearth)
require(tidyverse)
library(lubridate)
```

# Function definitions

Helper function to derive data via the ohsome API and return them as *sf* objects.

```{r}
getOsomeStat <- function(uri, valueFieldName="value", ...)
{
  results <- postForm(uri, binary=FALSE, .encoding = "utf-8", ...)
  
  #browser()
  resultList <- RJSONIO::fromJSON(results, simplify = TRUE)
  resultsDf <- data.frame(do.call("rbind", (resultList$result)))
  # make sure the right data types are used
  # for users we have  fromTimestamp  and  toTimestamp fields, not timestampe
  if(length(grep(x=names(resultsDf), pattern = "timestamp"))> 0)
  {
    resultsDf$timestamp <- parse_datetime( as.character(resultsDf$timestamp))  
  }
  if(length(grep(x=names(resultsDf), pattern = "fromTimestamp"))> 0)
  {
    resultsDf$fromTimestamp <- parse_datetime( as.character(resultsDf$fromTimestamp))  
  }
  if(length(grep(x=names(resultsDf), pattern = "toTimestamp"))> 0)
  {
    resultsDf$toTimestamp <- parse_datetime( as.character(resultsDf$toTimestamp))  
  }
  resultsDf[valueFieldName] <- parse_number( as.character(resultsDf$value))
  
  return(resultsDf)
}
```
# Load shapefile with admin boundaries

```{r, include=FALSE}
kreisSf <- st_read(dsn ="../data/VG250_shapes", layer = "VG250_KRS") kreisSf$SN_L <- as.numeric(as.character(kreisSf$SN_L))
kreisSf <- filter(kreisSf,SN_L %in% c("8", "14") )
state <- as.character(c("Baden-Württemberg", "Saxony"))
SN_L <- as.numeric(c(8, 14))
df <- data.frame(state, SN_L)
kreisSf <- full_join(kreisSf, df, by="SN_L")
kreisSf$RS <- as.numeric(as.character(kreisSf$RS))

districtTypes <- read.csv2("../data/districtTypes2017inkar.csv")
districtTypes$RS <- as.numeric(as.character(districtTypes$RS))
districtTypes $districtTypeNr <- as.numeric(as.character(districtTypes$districtTypeNr))
kreisSf <- left_join(kreisSf, districtTypes , by="RS")
districtTypes$RS <- as.numeric(as.character(districtTypes$RS))

kreisSf$area <- st_area(kreisSf)
```

Data need to be transformed into geographic coordinates

```{r, include=FALSE}
kreisSfwgs84 <-st_transform(kreisSf, crs = 4326)
```

```{r, include=FALSE}
tm_shape(kreisSfwgs84) + tm_borders()
```

Some error checking... only Konstanz consists of two rows with the same RS_0

```{r, include=FALSE}
#View(kreisSf_BW)
length(unique(kreisSf$RS_0))
nrow(kreisSf)
xtabs(~ RS_0, data=kreisSf )
idx <- which(kreisSf$RS_0 == "083350000000")
kreisSf[idx,]
```


Dissolve polygons based on Kreisname (for Konstanz, Leipzig, Karlsruhe and Heilbronn there were several Polygons) - Kreisname (GEN) is problematic since not unique e.g. for Heilbronn, Karlsruhe, Leipzig... 

```{r, include=FALSE}
# history: Used GEN before, now switched to RS_0
kreisSfwgs84_dissolved <- kreisSfwgs84 %>% group_by(RS_0) %>% summarize(area= sum(area))
```

Join info on GEN and state

```{r, include=FALSE}
addOnInfo <- st_drop_geometry(kreisSfwgs84) %>% select(GEN, RS_0, state, BEZ,  NUTS, districtType)
kreisSfwgs84_dissolved <- merge(kreisSfwgs84_dissolved, addOnInfo, by="RS_0")
```

Test plot of dissolved districts.

```{r}
tm_shape(kreisSfwgs84_dissolved) + tm_polygons(col="districtType")
```


# Get count statistics via ohsome

The following values for the key shop were used.

"shop" - values representing retail stores: "alcohol", "bakery", "beverages, butcher", "cheese", "chocolate", "coffee", "confectionery", "convenience", "dairy", "deli", "frozen_food", "greengrocer", "health_food", "ice_cream", "organic", "pasta", "pastry", "seafood", "spices", "tea", "wine", "water", "department_store", "general", "kiosk", "mall", "supermarket", "baby_goods", "bag", "boutique", "clothes", "fabric", "fashion_accessories", "jewelry", "leather", "sewing", "shoes", "tailor", "watches", "charity", "second_hand", "variety_store", "chemist", "cosmetics", "erotic", "hairdresser_supply", "hearing_aids", "herbalist", "medical_supply", "nutrition_supplements", "optician", "perfumery", "agrarian", "appliance", "bathroom_furnishing", "doityourself", "electrical", "energy", "fireplace", "florist", "garden_centre", "garden_furniture", "gas", "hardware", "houseware", "paint", "security", "trade", "windows", "antiques", "bed", "candles", "carpet", "curtain", "doors", "flooring", "furniture", "interior_decoration", "kitchen, lamps", "tiles", "window_blind", "computer", "robot", "electronics", "hifi", "mobile_phone", "radiotechnics", "vacuum_cleaner", "Atv", "bicycle", "boat", "car", "car_parts", "caravan", "fuel", "fishing", "free_flying", "golf", "hunting", "jetski", "motorcycle", "outdoor", "scuba_diving", "ski", "snowmobile", "sports", "swimming_pool", "trailer", "tyres", "art", "collector", "craft", "frame", "games", "model", "music", "musical_instrument", "camera", "trophy", "video", "video_games", "anime", "books", "gift", "lottery", "newsagent", "stationery", "ticket", "cannabis", "e-cigarette", "funeral_directors", "laundry", "party", "pet", "pest_control", "pest_control", "pyrotechnics", "religion", "tobacco", "toys", "travel_agency", "weapons", "outpost", "yes", "beauty", "car_repair", "copyshop", "dry_cleaning", "farm", "glaziery", "hairdresser", "locksmith", "massage", "pet_grooming", "photo", "tailor", "tattoo"

"shop" - values representing no retail stores: "bookmaker", "money_lender", "pawnbroker", "storage_rental", "vacant"

Query data for each district and store results in  data.frame.

```{r, warning=FALSE, message= FALSE}
 theValues <- c("bookmaker", "money_lender", "pawnbroker", "storage_rental", "vacant")
shopData <- NULL
for(i in 1:nrow(kreisSfwgs84_dissolved))
{
  aKreisSf <- kreisSfwgs84_dissolved[i,]
  aKreisGeoJ <- sf_geojson(aKreisSf)
  res <- getOsomeStat(uri = "https://api.ohsome.org/v1/elements/count/", bpolys = aKreisGeoJ, keys = "shop", time = "2008-01-01/2020-01-01/P1M", types="node", valueFieldName = "countShopTotal")
  res$district<- aKreisSf$GEN.x
  res$RS<- aKreisSf$RS_0
  res$state <- aKreisSf$state.x
  res$area <- aKreisSf$area
  res$NUTS <- aKreisSf$NUTS.x
  #loop over the values and count the number of elements for that shop
  # also add them to the res data.frame/tibble
  for(aValue in theValues)
  {
    valueFieldName= paste0("count_", aValue)
    resSingelShop <- getOsomeStat(uri = "https://api.ohsome.org/v1/elements/count/", bpolys = aKreisGeoJ, keys = "shop", time = "2008-01-01/2020-01-01/P1M", types="node", values = aValue, valueFieldName = valueFieldName)
    res[,valueFieldName] <- resSingelShop[,valueFieldName]
  }
  if(is.null(shopData))
  {
    shopData <- res
  } else {
    shopData <- dplyr::bind_rows(shopData, res)  
  }
  
}

head(shopData)

```

For multi-polygons numbers would need to be added - this is not necessary here  since multi-polygones have been merged before).

"amenity" - values representing retail stores:"fuel", "pharmacy"

```{r, warning=FALSE, message= FALSE}
 theValues <- c("fuel", "pharmacy")
amenityData <- NULL
for(i in 1:nrow(kreisSfwgs84_dissolved))
{
  aKreisSf <- kreisSfwgs84_dissolved[i,]
  aKreisGeoJ <- sf_geojson(aKreisSf)
  res <- getOsomeStat(uri = "https://api.ohsome.org/v1/elements/count/", bpolys = aKreisGeoJ, keys = "amenity", time = "2008-01-01/2020-01-01/P1M", types="node", valueFieldName = "countamenityTotal")
  res$district<- aKreisSf$GEN.x
  res$RS<- aKreisSf$RS_0
  res$state <- aKreisSf$state.x
  res$area <- aKreisSf$area
  res$NUTS <- aKreisSf$NUTS.x
  #loop over the values and count the number of elements for that amenity  # also add them to the res data.frame/tibble
  for(aValue in theValues)
  {
    valueFieldName= paste0("count_", aValue)
    resSingelamenity <- getOsomeStat(uri = "https://api.ohsome.org/v1/elements/count/", bpolys = aKreisGeoJ, keys = "amenity", time = "2008-01-01/2020-01-01/P1M", types="node", values = aValue, valueFieldName = valueFieldName)
    res[,valueFieldName] <- resSingelamenity[,valueFieldName]
  }
  if(is.null(amenityData))
  {
    amenityData <- res
  } else {
    amenityData <- dplyr::bind_rows(amenityData, res)  
  }
  
}

head(amenityData)

```


Get number of retail stores:
We compute the number of retail stores tagged as "shop" by subtraction of the number of shops, that representing no retail stores from the total number of shops.
The total number of retail stores is the sum of retail stores tagged "amenity" and of retail stores tagged "shop".

```{r, warning=FALSE, message= FALSE}
amenitySelect <- select(amenityData, count_fuel, count_pharmacy, RS, timestamp)
retailData <- dplyr::left_join(shopData, amenitySelect, by = c("RS", "timestamp"))
retailData <- mutate(retailData, retailStores = countShopTotal - count_bookmaker - count_money_lender - count_pawnbroker - count_storage_rental - count_vacant +count_fuel + count_pharmacy)

retailData <- retailData %>% select(timestamp, district, RS, state, area, NUTS, retailStores)

View(retailData)
save(retailData, file = "../data/retailData.Rdata")
write.table(retailData, file = "../data/retailData.csv", sep=";", row.names = FALSE)
```