---
title: "Companion to Assessing shop completeness in OpenStreetMap for two federalstates in Germany - Analysis of driving factors for shop completeness"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

```{r setup, include=FALSE}
# the used output format also requires that the rmdformats package has been installed
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(MASS)
library(ggplot2)
#library(VGAM)
#library(polycor)
library(plotly)
library(tidyverse)
#library(stats)
library(car)
require(ggeffects)
```


```{r, include=FALSE}
TabFactors <- read.csv2("../data/TabFactors.csv")




TabFactors$completeness <- as.numeric(as.character(TabFactors$completeness))

# define district type factor and factor levels
TabFactors$districtTypeF <- factor(TabFactors$districtType, levels=c(3,2,1), labels = c("rural districts", "urban districts", "independent cities"))

TabFactorsAll <- TabFactors
TabFactorsAll$noSaturationEstimated <- is.na(TabFactorsAll$asymptote)

# delete rows, asymptote = NA
TabFactors <- TabFactors[-c(which(is.na(TabFactors$asymptote))),]

```

```{r, include=FALSE}
TabFactors$district <- gsub(pattern="\xf6", replacement = "ö", TabFactors$district)
TabFactors$district <- gsub(pattern="\xe4", replacement = "ä", TabFactors$district)
TabFactors$district <- gsub(pattern="\xdf", replacement = "ß", TabFactors$district)
TabFactors$district <- gsub(pattern="\xfc", replacement = "ü", TabFactors$district)
```

The data set used for the analysis contains the following columns:
  
  - populationDensity: inhabitants per km² (for 2018)
  - avShops: current number of retail stores as average value from 2018 to 2020
  - asymptote: estimated asymptote via nonlinear regression
  - completeness: estimated completeness level [%]
  - func: best fit function (SSlogis = three parameter logistic function, SSfpl = four parameter logistic, SSasymp = asymptotic function, SSmicmen = Michaelis-Menten function)
  - district type: 
    - 1 = independent city
    - 2 = urban district
    - 3 = rural district
  - unemployments: rate of unemployed in the civilian labor force [%] (for 2017)
  - academics: employees to social security contributions at the place of residence with an academic qualification per 100 inhabitants of working age (for 2017)
  - GDP: the gross domestic product (GDP) per person in employment [1000 €] (for 2016)

# Explorative analysis

```{r}
# diagram shops and estimated asymptote
p1 <- ggplot(TabFactors, aes(x=avShops, y=asymptote, text=district)) + theme_minimal()  + labs(title="number of shops and asymptote") + ylab("asymptote") + xlab("number of shops") + 
  geom_point(lwd=0.5, alpha=0.8, aes(colour=districtTypeF)) +
  geom_abline(slope=1, intercept = 0, col="grey", lty=2) +
  theme(plot.title = element_text(size=10, face = "bold"))+ theme(axis.title.y=element_text(size=9)) +theme(axis.title.x=element_text(size=9))+ theme(legend.text = element_text(size = 9)) + theme(axis.text.y =  element_text(size = 7), axis.text.x =  element_text(size =7)) +
  #theme(legend.position = "bottom") +
  theme(legend.title = element_text(size = 9, face = "bold")) 

ggplotly(p1, Tooltip=c("district"))

#violin plot
ggplot(TabFactors, aes(x=districtTypeF, y=completeness, colour = districtTypeF)) + geom_violin(size=0.3) + 
  scale_fill_brewer(palette="Dark2") + geom_boxplot(width=0.1, size = 0.3)+
  #stat_summary(fun.y=median, geom="point", size=2, color="red")+
  #stat_summary(fun.y=mean, geom="point", shape=23, size=2)+
  labs(title = "completeness distribution by district type", y = "completeness level [%]", x="")+ theme(legend.position = "none") +  theme(plot.title = element_text(size=10, face="bold")) + theme(axis.title.y=element_text(size=9)) + theme(legend.text = element_text(size = 9)) + theme(axis.text.y =  element_text(size = 7), axis.text.x =  element_text(size =9, colour="black"))+
  geom_dotplot(binaxis='y', stackdir='center',position=position_dodge(1), binwidth=0.5, colour = "Black")
```

For interpretation on might want to take into account that more urban areas have also a higher unemployment rate:

```{r}
ggplot(TabFactors, aes(x=districtTypeF, y= unemployments)) + geom_violin()
```


```{r}
gUenmp <- lm(unemployments ~ districtTypeF, data= TabFactors)
summary(gUenmp)
```

However, differences are not strongly significant.


# Regression analysis 

A poisson GLM estimates significant coefficients for all four predictors considered in the analysis:

```{r}
g <- glm(asymptote ~ districtTypeF + GDP + unemployments + academics + offset(log(avShops)),  data=TabFactors, family=poisson)

## deviance, estimates and significance
1-g$deviance/g$null.deviance
summary(g)

## VIF

vif(g)
```

## Adjust for overdispersion

However, to account for overdispersion in the data we have to use a negative binomial GLM. This reduces the number of significant predictors due to the additional uncertainty that was (wrongly) ignored by the poisson GLM.

```{r}
g <- glm.nb(asymptote ~ districtTypeF + GDP + unemployments + academics + offset(log(avShops)),  data=TabFactors)

## deviance, estimates and significance
1-g$deviance/g$null.deviance
summary(g)

## VIF

vif(g)
```

Need to simplify model:

```{r}
g <- update(g, ~. -academics)
summary(g)
vif(g)
```
```{r}
drop1(g)
```
(This is the model used in the manuscript.)

```{r}
g <- update(g, ~. -GDP)
summary(g)
vif(g)
1 - g$deviance / g$null.deviance
```



Compare with model with GDP instead of unemployment rate

```{r}
g2 <- glm.nb(formula = asymptote ~ districtTypeF + GDP + offset(log(avShops)), data = TabFactors, link = log)
summary(g2)
```
Clearly worst, so we stick with g. Are were interactions?

```{r}
g2 <- glm.nb(formula = asymptote ~ districtTypeF * unemployments + offset(log(avShops)), data = TabFactors, link = log)
summary(g2)
```
Nope.




### Diagnostics
```{r}
plot(g, which=1:6, labels.id = TabFactors$district)
```

Nordsachsen is highly influential, characterized by a high cook's distance and a high leverage. Görlitz is also remarkable due to its high leverage.

### What if we leave out Nordsachsen?

```{r}
#TabFactors$district
gPart <- glm.nb(asymptote ~ unemployments +  offset(log(avShops)),  data=TabFactors, subset= district != "Nordsachsen, rural district" )

## deviance, estimates and significance
1-gPart$deviance/gPart$null.deviance
summary(gPart)
```
If we drop Görlitz as another district with high leverage, the relationship with unemployment holds. 

```{r}
gPart <- glm.nb(asymptote ~ unemployments +  offset(log(avShops)),  data=TabFactors, subset= district != "Nordsachsen, rural district" | district != "Görlitz, rural district")

## deviance, estimates and significance
1-gPart$deviance/gPart$null.deviance
summary(gPart)
```
If we only drop Görlitz 

```{r}
gPart <- glm.nb(asymptote ~ districtTypeF + unemployments +  offset(log(avShops)),  data=TabFactors, subset= district != "Nordsachsen, rural district" | district != "Görlitz, rural district")

## deviance, estimates and significance
1-gPart$deviance/gPart$null.deviance
summary(gPart)
```

Coefficient estimates stay the same if Görlitz is excluded or not.


# Districts for those no reliable saturation level could be estimated


```{r}
gSat <- glm(noSaturationEstimated ~ districtTypeF + unemployments + GDP, data=TabFactorsAll, family=binomial)
summary(gSat)
drop1(gSat)
gSat <- update(gSat, ~.- districtTypeF)
summary(gSat)
drop1(gSat)
gSat <- update(gSat, ~.- unemployments)
summary(gSat)
drop1(gSat)

```





```{r}
ggplot(TabFactorsAll, aes(x=noSaturationEstimated, y= GDP)) + geom_boxplot(varwidth=TRUE) + facet_wrap(~districtTypeF)
```


```{r}
ggplot(TabFactorsAll, aes(x=noSaturationEstimated, y= unemployments)) + geom_boxplot(varwidth=TRUE) + facet_wrap(~districtTypeF)
```
