---
title: "Companion to Assessing shop completeness in OpenStreetMap for two federalstates in Germany - non linear regression of OSM retail stores time series by varied selfstarting functions"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
require(ggplot2)
require(stats)
require(nlstools)
require(tidyverse)
require(dplyr)
```


```{r, include=FALSE}
load("../data/retailData.Rdata")
```


```{r, include=FALSE}
RS <- as.character(c())
district <- as.character(c())
state <- as.character(c())
avStores <- as.numeric(c()) # current number of OSM retail stores (average value during the last 2 years of the analysis period)

#the three parameter logistic function:
SSlogis_asymptote <- as.numeric(c()) # estimate asymptote
SSlogis_Std.Error <- as.numeric(c()) # standard error asymptote
SSlogis_ResErrorRel <- as.numeric(c()) #relativ residual standard error
SSlogis_SlopeDev <- as.numeric(c()) #Slope Deviation

#four parameter logistic function: 
SSfpl_asymptote <- as.numeric(c()) # estimate asymptote
SSfpl_Std.Error <- as.numeric(c()) # standard error asymptote
SSfpl_ResErrorRel <- as.numeric(c()) #relativ residual standard error
SSfpl_SlopeDev <- as.numeric(c()) #Slope Deviation

#asymptotic function:
SSasymp_asymptote <- as.numeric(c()) # estimate asymptote
SSasymp_Std.Error <- as.numeric(c()) # standard error asymptote
SSasymp_ResErrorRel <- as.numeric(c()) #relativ residual standard error
SSasymp_SlopeDev <- as.numeric(c()) #Slope Deviation

#rectangular hyperbola (Michaelis Menten) function:
SSmicmen_asymptote <- as.numeric(c()) # estimate asymptote
SSmicmen_Std.Error <- as.numeric(c()) # standard error asymptote
SSmicmen_ResErrorRel <- as.numeric(c()) #relativ residual standard error
SSmicmen_SlopeDev <- as.numeric(c()) #Slope Deviation

nlsData <- data.frame(RS, district, state, avStores, SSlogis_asymptote, SSlogis_Std.Error, SSlogis_ResErrorRel, SSlogis_SlopeDev, SSfpl_asymptote, SSfpl_Std.Error, SSfpl_ResErrorRel, SSfpl_SlopeDev, SSasymp_asymptote, SSasymp_Std.Error, SSasymp_ResErrorRel, SSasymp_SlopeDev, SSmicmen_asymptote, SSmicmen_Std.Error, SSmicmen_ResErrorRel, SSmicmen_SlopeDev)
```


```{r, include=FALSE}
StartDate <- ISOdate(2008,01,01,00) # start analysis period
class(StartDate)
intervall <- retailData$timestamp - StartDate 
retailData$time <- as.numeric(intervall, units="days") + 1
# + 1, set retailData$time[1] = 1 instead of 0
head(retailData)
```

Create dataframe including criteria of all fits and further information:

table parameter:

avStores = current number of OSM retail stores (average value during the last 2 years of the analysis period)

asymptote = estimated asymptote by nls approach

Std.Error = standard error of the estimated asymptote

ResErrorRel = relativ residual standard error (normalized by avStores)

SlopeDev = relative deviation of the slope between the historic development and the fitted curve during the last 2 years of the analysis period


```{r}
RSselect <- as.character('')

for(i in 1:nrow(retailData))
{ 
  if (RSselect != retailData$RS[i])
  {
    RSselect <- as.character(retailData$RS[i])
    state <- as.character(retailData$state[i])
    district <- as.character(retailData$district[i])
    RS <- RSselect
    districtData <- dplyr::filter(retailData, RS %in% RSselect)
    avStores <- as.numeric(round(sum(districtData$retailStores[122:145])/24))
    
    
# non-linear least squares by the three parameter logistic function SSlogis:
    LogisModel <- nls(retailStores ~ SSlogis(time, Asym, xmid, scal), districtData)
    #summary(LogisModel)
    Asym <- (summary(LogisModel)$coefficients[1, 1]) 
    xmid <- (summary(LogisModel)$coefficients[2, 1]) 
    scal <- (summary(LogisModel)$coefficients[3, 1])
    

    SSlogis_asymptote <- as.numeric(round(summary(LogisModel)$coefficients[1, 1]))
    SSlogis_Std.Error <- as.numeric(round(summary(LogisModel)$coefficients[1, 2]))
    SSlogis_ResErrorRel <- as.numeric(round((summary(LogisModel)$sigma)/districtData$retailStores[145], digits = 2))
    Asym <- (summary(LogisModel)$coefficients[1, 1]) #asymptote
    xmid <- (summary(LogisModel)$coefficients[2, 1]) 
    scal <- (summary(LogisModel)$coefficients[3, 1])
    
    
       LogisFunc<-function(x){
       y <- Asym/(1+exp((xmid-x)/scal))   
       return(y)
       } 
          
    # slope of the fit curve durging the last years of the analysis period
    slopeSSlogis <- as.numeric(round((LogisFunc(4384) - LogisFunc(3685))/2))
    
   # slope of the historic development durging the last years of the analysis period 
    slopeData <- as.numeric(round((districtData$retailStores[145]-districtData$retailStores[122])/2))
    
    SSlogis_SlopeDev <- slopeSSlogis - slopeData
    
    
# non-linear least squares by the four parameter logistic function SSfpl: 
   FplModel  <- tryCatch(nls(retailStores ~ SSfpl(time, A, B, xmid, scal), districtData),
    error = function(e) paste0("NA"))
    
    if(FplModel[1] %in% "NA")
    {    # add to dataframe
    SSfpl_asymptote <- NA
    SSfpl_Std.Error <- NA
    SSfpl_ResErrorRel <- NA
    SSfpl_SlopeDev <- NA
      } else {
    A <- (summary(FplModel)$coefficients[1, 1]) 
    B <- (summary(FplModel)$coefficients[2, 1]) #asymptote
    xmid <- (summary(FplModel)$coefficients[3, 1])
    scal <- (summary(FplModel)$coefficients[4, 1])
        
    SSfpl_asymptote <- as.numeric(round(summary(FplModel)$coefficients[2, 1]))
    SSfpl_Std.Error <- as.numeric(round(summary(FplModel)$coefficients[2, 2]))
    SSfpl_ResErrorRel <- as.numeric(round((summary(FplModel)$sigma)/districtData$retailStores[145], digits = 2))

      FplFunc<-function(x){
      y <- A+(B-A)/(1+exp((xmid-x)/scal))   
      return(y)
    } 
          
    # slope of the fit curve durging the last years of the analysis period
    slopeSSfpl <- as.numeric(round((FplFunc(4384) - FplFunc(3685))/2))
    
    # deviation of the slope of the real data and the fit curve via SSfpl
    SSfpl_SlopeDev <- slopeSSfpl - slopeData
      }   
 
   # non-linear least squares by the asymptotic function SSasymp:
    AsympModel <- tryCatch(nls(retailStores ~ SSasymp(time, Asym, R0, lrc), data=districtData),  error = function(e) paste0("NA"))
    
    if(AsympModel[1] %in% "NA")
    {   
    SSasymp_asymptote <- NA
    SSasymp_Std.Error <- NA
    SSasymp_ResErrorRel <- NA
    SSasymp_SlopeDev <- NA
      } else {
    Asym <- (summary(AsympModel)$coefficients[1, 1]) #Asymptote
    R0 <- (summary(AsympModel)$coefficients[2, 1])
    lrc <- (summary(AsympModel)$coefficients[3, 1])

    SSasymp_asymptote <- as.numeric(round(summary(AsympModel)$coefficients[1, 1]))
    SSasymp_Std.Error <- as.numeric(round(summary(AsympModel)$coefficients[1, 2]))
    SSasymp_ResErrorRel <- as.numeric(round((summary(AsympModel)$sigma)/districtData$retailStores[145], digits = 2))

    AsympFunc<-function(x){
      y <- Asym+(R0-Asym)*exp(-exp(lrc)*x)   
      return(y)
    } 
          
    # slope of the fit curve durging the last years of the analysis period
    slopeSSasymp <- as.numeric(round((AsympFunc(4384) - AsympFunc(3685))/2))
    
    
    # deviation of the slope of the real data and the fit curve via SSasymp
    SSasymp_SlopeDev <- slopeSSasymp - slopeData
      }   
    
    
       # non-linear least squares by the rectangular hyperbola (Michaelis Menten) function SSmicmen
      MicmenModel  <- tryCatch(nls(retailStores ~ SSmicmen(time, Vm, K), data=districtData),
    error = function(e) paste0("NA"))
    
    if(MicmenModel[1] %in% "NA")
    {   
    SSmicmen_asymptote <- NA
    SSmicmen_Std.Error <- NA
    SSmicmen_ResErrorRel <- NA
    SSmicmen_SlopeDev <- NA
      } else {
    Vm <- (summary(MicmenModel)$coefficients[1, 1]) #Asymptote
    K <- (summary(MicmenModel)$coefficients[2, 1])

    SSmicmen_asymptote <- as.numeric(round(summary(MicmenModel)$coefficients[1, 1]))
    SSmicmen_Std.Error <- as.numeric(round(summary(MicmenModel)$coefficients[1, 2]))
    SSmicmen_ResErrorRel <- as.numeric(round((summary(MicmenModel)$sigma)/districtData$retailStores[145], digits = 2))

     MicmenFunc<-function(x){
      y <- Vm*x/(K+x)   
      return(y)
    } 
          
    # slope of the fit curve during the last years of the analysis period
    slopeSSmicmen <- as.numeric(round((MicmenFunc(4384) - MicmenFunc(3685))/2))
    
    
    # deviation of the slope of the real data and the fit curve via SSmicmen
    SSmicmen_SlopeDev <- slopeSSmicmen - slopeData
      }    
      
    newFrame <- data.frame(RS, district, state, avStores, SSlogis_asymptote, SSlogis_Std.Error, SSlogis_ResErrorRel, SSlogis_SlopeDev, SSfpl_asymptote, SSfpl_Std.Error, SSfpl_ResErrorRel, SSfpl_SlopeDev, SSasymp_asymptote, SSasymp_Std.Error, SSasymp_ResErrorRel, SSasymp_SlopeDev, SSmicmen_asymptote, SSmicmen_Std.Error, SSmicmen_ResErrorRel, SSmicmen_SlopeDev)
    
    nlsData <- dplyr::bind_rows(nlsData, newFrame)
  } }

```



```{r}
print(nlsData)
```

A few overview plots about the results

```{r, fig.height=9}
ggplot(nlsData, aes(x=district, y=SSlogis_asymptote)) + geom_point() + coord_flip()
```

```{r, fig.height=9}
ggplot(nlsData, aes(x=district, y=SSfpl_asymptote)) + geom_point() + coord_flip()
```

```{r, fig.height=9}
ggplot(nlsData, aes(x=district, y=SSasymp_asymptote)) + geom_point() + coord_flip()
```

```{r, fig.height=9}
ggplot(nlsData, aes(x=district, y=SSmicmen_asymptote)) + geom_point() + coord_flip() 
```

```{r, fig.height=9}
ggplot(nlsData, aes(x=district, y=avStores)) + geom_point() + coord_flip() 
```

```{r, include=FALSE}
# export and serialize results
save(nlsData, file = "nlsData.Rdata")
write.table(nlsData, file = "nlsData.csv",sep="", dec=".", row.names = FALSE)
```
